Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-04 paper/main.tex
paper s1-2

\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}

%% Revision documents
%\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% FIXME: \sout is not very good. It cannot handle commands within its argument.
\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
\let\svthefootnote\thefootnote
\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
% Command for edits, command for commenting and color
\newcommand\newrevisor[3]{%%
  \colorlet{#1}{#3}
  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
}
\newrevisor{manuel}{MANUEL}{Purple}
\newrevisor{ekhine}{EKHINE}{red}
%\usepackage{showframe}

@@

\EKHINE{6. When the main contribution of the paper is discussed it is stated that this contribution is the Unbalanced Mallows *model* (which does not seem to be true as the models used in the paper seem to be known already). Right after that it is stated that the contribution is *a novel algorithm* which, I think, reflects the contents of the paper much better.
}
In this work, we propose and analyze Unbalanced Mallows Model (UMM)\MANUEL{Perhaps we need to rename the algorithm to UMM Algorithm or UMM Optimizer to avoid the criticism of the reviewer that we do not propose the model?}, a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There
exists a wide body of research on probabilistic models for permutations, such
as Mallows~\cite{FliVer1986}, that have already been  used in optimization
problems successfully, with the most prominent examples being estimation of
distribution algorithms
(EDAs)~\cite{CebIruMen2014eda}. ACO may also be considered
a type of EDA, since it builds a probability distribution model from which
solutions are sampled with a bias towards the best solutions evaluated so
far. These classical EDAs  assume that function
evaluations are cheap and/or the fitness function is white-box.%\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
%The proposed UMM drastically differs from the above algorithms.

@@

An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}. \EKHINE{ojo, aqui lo dice bien o  mal?}\MANUEL{Habría que mirarlo bien!}

@@

Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off. At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$ (half of the expected distance for the uniform distribution),\MANUEL{Reviewer 3: "At the first iteration, $\theta$ is set so that $E[D] = {n \choose 4}$". Why?} and then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$. Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.

\EKHINE{5. "At the first iteration, Theta is set so that $E[D] = {n \choose 4}$". Why?}
A purely explotation approach would always generate $\sigma_0$ in the sampling stage. However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs.  Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.

@@

\ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
 obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
instance considered here. The lack of any apparent convergence suggests that
the models (both Bayesian and probabilistic) are not learning anything about
the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{}
% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
% that other distance measures are more suited than Kendall's-$\tau$ distance for
% the PFSP, the reported fitness differences between various distance metrics are
% small.
% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}

\begin{figure}[tb]
  \centering%
  \includegraphics[width=\textwidth]{../img/fitness_real_pfsp_rec05_txt}
  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure}

>>>>

\newcommand{\supplement}{\href{http://doi.org/10.5281/zenodo.4500974}{doi:~10.5281/zenodo.4500974}}
\newcommand{\ken}{Kendall's-$\tau$}
%% Revision documents
%\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
% FIXME: \sout is not very good. It cannot handle commands within its argument.
\newcommand{\SuggestEdit}[3][red]{\textcolor{#1}{\sout{#2}}\textcolor{#1}{#3}}
\let\svthefootnote\thefootnote
\newcommand\colorfootnote[2][black]{\def\thefootnote{\color{#1}\svthefootnote}%
  \footnote{\color{#1}#2}\def\thefootnote{\color{black}\svthefootnote}}
\newcommand\RevComment[3][red]{\protect\colorfootnote[#1]{{\textbf{[#2: #3]}}}}
% Command for edits, command for commenting and color
\newcommand\newrevisor[3]{%%
  \colorlet{#1}{#3}
  \expandafter\newcommand\csname#1\endcsname[2]{\SuggestEdit[#1]{##1}{##2}}%%
  \expandafter\newcommand\csname#2\endcsname[1]{\RevComment[#1]{#2}{##1}}%%
}
\newrevisor{manuel}{MANUEL}{Purple}
\newrevisor{ekhine}{EKHINE}{red}
%\usepackage{showframe}

@@

In this work, we propose and analyze the Unbalanced Mallows Model (UMM) algorithm, a
population-based probabilistic algorithm specifically designed for optimizing
black-box problems over a permutation landscape under a limited budget. There\EKHINE{En realidad está disenado para esto  pero podríamos probarlo para otros casos. Trabajo futuro}
exists a wide body of research on probabilistic models for permutations, such
as Mallows~\cite{FliVer1986}, that have already been  used in optimization
problems successfully, with the most prominent examples being estimation of
distribution algorithms
(EDAs)~\cite{CebIruMen2014eda}. ACO may also be considered
a type of EDA, since it builds a probability distribution model from which
solutions are sampled with a bias towards the best solutions evaluated so
far. These classical EDAs  assume that function
evaluations are cheap and/or the fitness function is white-box.%\MANUEL{Do EDAs really assume that the fitness function is white-box? ACO doesn't: They are hybridized with local search }
%The proposed UMM drastically differs from the above algorithms.

@@

An equivalent definition for the Kendall's-$\tau$ distance counts the number of adjacent swaps that have to be made to convert $\sigma_1^{-1}$ into $\sigma_2^{-1}$ and, therefore, it is sometimes called \emph{swap distance}~\citep{ZaeStoBar2014:ppsn}. \EKHINE{Mirandolo bien, lo que pasa es que dicen que es lo mismo la swap distance que la kendall!}

@@

Parameter $\theta$ is set automatically by the algorithm to control the diversification and intensification trade off. At the first iteration, $\theta$ is set so that  $\Edist = \binom{n}{4}$, i.e., half of the expected distance for the uniform distribution. 
Then the expected distance linearly decreases until the last iteration where $\theta$ is set such that $\Edist = 1$. Despite it is not possible to isolate $\theta$ in Eq.~\eqref{eq:expectation}, its monotonicity allows the use of bisection methods for efficiently finding an approximation.

Setting $\Edist = \binom{n}{2}$ will generate random permutations at early stages, slowing the algorithm convergence. 
A purely explotation approach will set $\Edist = 0$  generating $\sigma_0$ in the sampling stage. However, the setting of a decreasing variance   controls the exploration-explotation tradeoff as a function of the number of function evaluations that the algorithm performs. It follows that the behaviour is different for different choices of $m$. Similar approaches have been taken in different contexts~\cite{ArzCebPer2019qap}.

@@

\ekhine{In the case of the PFSP, the progress of both algorithms resembles a random search, as illustrated by the results
 obtained on \texttt{rec05} (Fig.~\ref{fig:rec05}), which is the smallest PFSP
instance considered here. The lack of any apparent convergence suggests that
the models (both Bayesian and probabilistic) are not learning anything about
the fitness landscape. Results for larger PFSP corroborate these conclusions (\supplement).}{kaka}
% \manuel{Although previous studies~\citep{ZaeStoBar2014:ppsn} concluded
% that other distance measures are more suited than Kendall's-$\tau$ distance for
% the PFSP, the reported fitness differences between various distance metrics are
% small.
% Hence, we conjecture that the same behavior will be observed with other distance metrics.}{}\MANUEL{On second thought, I'm not sure we should say this, because we really have no idea and we actually hope it is not true.}

\begin{figure*}[tb]
  \centering%
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_IO_N-t59b11xx}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_SGB_N-sgb75_02}

    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p40-01}
    \includegraphics[width=0.49\textwidth]{../img/fitness_lop_RandB_N-p50-01}

    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec19_txt}
    \includegraphics[width=0.49\textwidth]{../img/fitness_pfsp_rec31_txt}

  \caption{Mean fitness  (and standard deviation)  of each solution evaluated by each algorithm on the PFSP instance \texttt{rec05}.\label{fig:rec05}}
\end{figure*}
